---
title: "MSCRCclassifier: an R package for multi-omics subtyping of colorectal cancer"
output:
  BiocStyle::html_document:
    toc: FALSE
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
date: ""
---

<style>
h1.title {
  text-align: center;
}
</style>

<!-- <div style="text-align: center;"> </div> -->
Lingli He<sup>1</sup> and Xin Wang<sup>1</sup>

<sup>1</sup> Department of Surgery, The Chinese University of Hong Kong, Hong Kong SAR, China.

* contact: xinwang@cuhk.edu.hk
* `r paste("date:", Sys.Date())`
* package: `r BiocStyle::pkg_ver("MSCRCclassifier")`


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

The vignette helps the user to do multi-omics colorectal cancer subtyping using sparse mCCA (@witten2009extensions) and weighted average. Piared mRNA expression, microRNA expression, and DNA methylation data from TCGA-COAD and TCGA-READ datasets were used for the training of multi-omics colorectal classifier. The package accepts any combination of mRNA expression, microRNA expression, and DNA methylation data as input. 


# Package installation

Please run all analyses in this vignette under version 2.10 of R Prior to installation of package MSCRCclassifier, R packages caret, naivebayes should be installed. These packages can be installed directly from CRAN (Comprehensive R Archive Network):
```{r message=FALSE, warning=FALSE, results='hide'}
options(repos = c(CRAN = "https://cloud.r-project.org/"))
install.packages(c("caret", "naivebayes"))
library(caret)
library(naivebayes)
library(devtools)
# install the "MSCRCclassifier" package
# install_github("Carpentierbio/MSCRCclassifier")
```

# Quick start 

## Projecting each omics data into unified space and inregraation
The example dataset used in this analysis is from a micro-array experiment on 566 colon cancer patients with GEO number named [GSE39582](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE39582)  (@marisa2013gene).

* input data can be any combination of mRNA expression, microRNA expression, and DNA methylation data.
* if the single-omics or multi-omics integrated data includes mRNA expression or miRNA expression, *log*~2~\ transformation is required.
* all types of input data should be pre-processed and z-score normalized.

```{r, eval=TRUE}
library(MSCRCclassifier)
library(dplyr)
# Load example mRNA expression profile
data("GSE39582_expr")
# Load projection matrixs derived from 315 TCGA-COAD and TCGA-READ samples
data("projection_mxs")
dim(projection_mxs$ws[[1]])  
# Projecting each omics data into unified space
mRNAexprCCA = t(GSE39582_expr) %*% projection_mxs$ws[[1]]
mRNAexprCCA = scale(mRNAexprCCA)
mRNAexprCCA[1:5,1:5] # samples in rows and genes in columns 算出来是行样本列基因
  
a1=0.4
data_input = scale(a1*mRNAexprCCA)
colnames(data_input) = paste("X",1:ncol(projection_mxs$ws[[1]]), sep = "")
data_input[1:5,1:5]
```

## Multi-omics colorectal cancer subtype classification
`classifyMSCRC` function requires an expression matrix with samples in rows and multi-omics features in columns as input. The column names of the expression profile should be X1, X2, ..., X196. The code chunk below demonstrates how to perform classification using primary colorectal cancer example data. 

```{r, eval=TRUE, message=TRUE}

# MSCRC prediction of primary colorectal cancer
result <- classifyMSCRC(data_input)
label <- result$label
prob <- result$prob %>%
  `colnames<-`(paste("MSCRC", 1:5, "_prob", sep = ""))
res = data.frame(prob, subtype = paste("MSCRC", label, sep = "") ) %>%
  `rownames<-`(names(label))
head(res)
```


# Session Info
```{r, echo=FALSE}
  sessionInfo()
```

# References



